SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-11-02 2017-Xudong-SMT/ActaInformatica.tex
some typos corrected

@@

Implementing these semantics raised several challenges:
\begin{itemize}
	\item to obtain a tool that could be applied to pNets representing
	various languages, in particular various action algebras,
	with their specific decision theories;
	\item to separate clearly the 
	algorithm generating the transitions of an open automaton from
	combination of all possible (symbolic) behaviours, and from
	the symbolic reasoning part, specifically here using an SMT
	engine to check the
	satisfiability of the predicates generated by our algorithm;
	\item to build a prototype of the algorithm, and validate the approach on our basic
	case-studies, meanwhile, understand the efficiency of the interaction
	with the SMT solver.
\end{itemize}

@@

From a broader perspective, after the seminal work
of Hennessy, Lin, and their colleagues on value-passing systems with
assignments, many
different works addressing various classes of infinite-state systems
and/or parameterised topologies have been published, using
combinations of approaches, often including predicate abstraction and
SMT satisfiability (\eg \cite{DBLP:journals/jsat/AlbertiGPRR12,BruniEtAl-Tiles-Concur2000,CimattiEtAl-NUXMV-CAV2014,ChampionEtAl-Kind2-CAV2016,DBLP:conf/cade/GhilardiNRZ08}). 
Among these works, several have shown the capability of the SMT engines
(either Z3 or Yikes) as servers for solving verification conditions of
 algorithms, for large case-studies
(\eg
\cite{DBLP:journals/corr/abs-1806-11459,DBLP:journals/corr/CimattiGMT13}).
With
respect to these, we use symbolic representations not only to get a
finite representation of infinite spaces, but also to express the
(data-sensitive) synchronisations with the environment, making our
models suitable for compositional verification.

@@

		\item[$\bullet$] $\symb{SV}_k^{k\in K}$ is a set of
		synchronisation vectors ($K\in\I_\P$ where $\I_\P$ is the set of indexed
		sets with a range depending on variables of $\P$).  It is required that each
		$\symb{SV}_k$ is a pair $(\alpha_{l}^{l\in I_k \uplus J_k}\to\alpha'_k,\ g_k)$, where
		$\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
			J_k}{\vars({\alpha_l})}$, $\alpha'_k\in \actions{\variables}$, $I_k\subseteq I$, and $J_k\subseteq J$. The global action of a vector $\symb{SV}_k$ is
		$\alpha'_k$. The Boolean expression $g_k $, with $\vars(g_k)\subseteq \bigcup_{l\in 
			I_k\uplus J_k}{\vars({\alpha_l})}$, is a guard associated to the vector.

@@

\vspace{-5mm}
\begin{figure}[ht]
\centerline{ 
  \includegraphics[width=0.25\linewidth]{ActaXFIG/Enable1-aut}
  \hspace{25mm}
  \includegraphics[width=0.20\linewidth]{ActaXFIG/Enable2-aut}}
  \begin{minipage}{5.7cm}
$  \begin{array}{@{}r@{}l@{}}
    ot_1  = &\openrule{\{ \xrightarrow{a_1}_P \},\\ [\forall y1. a_1\neq\delta(y1)]}
    {A_0 \xrightarrow{a_1} A_0}\\
    ot_2  =& \openrule{
                            \{\xrightarrow{\delta(x_1)}_P,\,
                            \xrightarrow{acc(x_1)}_Q\}  }
    {A_0 \xrightarrow{\underline{\delta(x_1)}} A_1}\\
    ot_3  =& \openrule{
                          \{  \xrightarrow{a_2}_Q \},\\ [\forall y2. a_2\neq acc(y2)] }
    {A_1 \xrightarrow{a_2} A_1}
  \end{array}$
  \end{minipage}
  \hspace{2mm}
  \begin{minipage}{6cm}
$  \begin{array}{@{}r@{}l@{}}
    ot'_1  =& \openrule{
    	\{\xrightarrow{b1}_P\}, ~~ [\forall y1. b1\neq\delta(y1) \!\land\! v=0]}
    {B_0 \xrightarrow{b1} B_0}\\
    ot'_2  =& \openrule{
                            \{\xrightarrow{\delta(y2)}_P,\,
                            \xrightarrow{acc(y2)}_Q \}, \\
                                        v=0, \\
                                        \{v\gets 1\}}
    {B_0 \xrightarrow{\underline{\delta(y2)}} B_1}\\
    ot'_3  =& \openrule{
                            \xrightarrow{b2}_q, \\ [\forall y2. b2\neq\delta(y2) \!\land\! v=1] }
    {B_1 \xrightarrow{b2} B_1}
  \end{array}$
  \end{minipage}
  \caption{The two open automata}
  \label{schema:enable-OAs}
\end{figure}

The open automata generated for the two encodings of Enable in
Figure~\ref{schema:enable-pnets} are displayed in
Figure~\ref{schema:enable-OAs}.
Note that although these automata look very similar to the PLTS
controllers from Figure~\ref{schema:enable-pnets}, their nature is very
different: they represent the behavior of the whole system, including
the holes $P$ and $Q$. Their transitions do not rely on a particular
implementation of the controllers, but only encode the 
relations between the actions of the holes.
We have shown in \cite{henrio:Forte2016} how to prove that they are bisimilar.

@@

\begin{figure}[h]
  \begin{minipage}{2cm}
  \includegraphics[width=0.9\linewidth]{ActaXFIG/PQR-automaton}
  \end{minipage}
  \hspace{10mm}
  \begin{minipage}{3.5cm}
  \begin{eqnarray*}
    OT_1 & = \openrule{\{\xrightarrow{a1}_P\}\\ a1\neq\delta(x)}
    {\ostate{00} \xrightarrow{a1} \ostate{00}}\\
    OT_2 & = \openrule{
                            \{\xrightarrow{\delta(x1)}_P, ~~
                            \xrightarrow{acc(x1)}_Q\}
                      }
                      {\ostate{00} \xrightarrow{\underline{\delta(x1)}} \ostate{10}}    \\
    OT_3 & = \openrule{
                        \{\xrightarrow{a2}_Q\},\\a2\neq\delta(x)
                        }
    {\ostate{10} \xrightarrow{a2} \ostate{10}}
    \end{eqnarray*}
  \end{minipage}
  \hspace{10mm}
  \begin{minipage}{4cm}
  \begin{eqnarray*}
    OT_4 & = \openrule{
                      \{\xrightarrow{\delta(x2)}_Q, \qquad 
                       \xrightarrow{acc(x2)}_R\} 
                       }
                      {\ostate{10} \xrightarrow{\underline{\delta(x2)}} \ostate{11}}    \\
    OT_5 & = \openrule{\{\xrightarrow{a3}_R\}
    	}
    {\ostate{11} \xrightarrow{a3} \ostate{11}}
    \end{eqnarray*}
  \end{minipage}
  \caption{Open automaton for the term ``$P>>(Q>>R)$''}
  \label{schema:enable3bis}
\end{figure}

>>>>

\newcommand{\YX}[1]{{\color{red}#1}}

@@

Implementing these semantics raised several challenges:
\begin{itemize}
	\item to obtain a tool that could be applied to pNets representing
	various languages, in particular various action algebras,
	with their specific decision theories;
	\item to separate clearly the 
	algorithm 
	\YX{that generates the transitions of an open automaton from
	the symbolic reasoning part that uses an SMT
	engine to check the
	satisfiability of the predicates generated by the algorithm;}

%	\item to separate clearly the algorithm generating the transitions of an open automaton from combination of all possible (symbolic) behaviours, and from the symbolic reasoning part, specifically here using an SMT engine to check the satisfiability of the predicates generated by our algorithm;

	\item to build a prototype of the algorithm, and validate the approach on our basic
	case-studies, meanwhile, understand the efficiency of the interaction
	with the SMT solver.
\end{itemize}

@@

From a broader perspective, after the seminal work
of Hennessy, Lin, and their colleagues on value-passing systems with
assignments, many
different works addressing various classes of infinite-state systems
and/or parameterised topologies have been published, using
combinations of approaches, often including predicate abstraction and
SMT satisfiability (\eg \cite{DBLP:journals/jsat/AlbertiGPRR12,BruniEtAl-Tiles-Concur2000,CimattiEtAl-NUXMV-CAV2014,ChampionEtAl-Kind2-CAV2016,DBLP:conf/cade/GhilardiNRZ08}). 
Among these works, several have % shown the capability 
\YX{
demonstrated the usefulness
of SMT engines
(either Z3 or Yikes) as servers for solving verification conditions of
 algorithms, especially for large case-studies
(\eg
\cite{DBLP:journals/corr/abs-1806-11459,DBLP:journals/corr/CimattiGMT13}).
}
With
respect to these, we use symbolic representations not only to get a
finite representation of infinite spaces, but also to express the
(data-sensitive) synchronisations with the environment, making our
models suitable for compositional verification.

@@

		\item[$\bullet$] $\symb{SV}_k^{k\in K}$ is a set of
		synchronisation vectors, \YX{where $K\in\I_\P$ with $\I_\P$ being the set of indexed
		sets whose ranges depend on the variables in $\P$}.  It is required that each
		$\symb{SV}_k$ is a pair $(\alpha_{l}^{l\in I_k \uplus J_k}\to\alpha'_k,\ g_k)$, where
		$\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
			J_k}{\vars({\alpha_l})}$, $\alpha'_k\in \actions{\variables}$, $I_k\subseteq I$, and $J_k\subseteq J$. The global action of a vector $\symb{SV}_k$ is
		$\alpha'_k$. The Boolean expression $g_k $, with $\vars(g_k)\subseteq \bigcup_{l\in 
			I_k\uplus J_k}{\vars({\alpha_l})}$, is a guard associated to the vector.

@@

\vspace{-5mm}
\begin{figure}[ht]
\centerline{ 
  \includegraphics[width=0.25\linewidth]{ActaXFIG/Enable1-aut}
  \hspace{25mm}
  \includegraphics[width=0.20\linewidth]{ActaXFIG/Enable2-aut}}
  \begin{minipage}{5.7cm}
$  \begin{array}{@{}r@{}l@{}}
    ot_1  = &\openrule{\{ \xrightarrow{a_1}_P \},\\ [\forall y1. a_1\neq\delta(y1)]}
    {A_0 \xrightarrow{a_1} A_0}\\
    ot_2  =& \openrule{
                            \{\xrightarrow{\delta(x_1)}_P,\,
                            \xrightarrow{acc(x_1)}_Q\}  }
    {A_0 \xrightarrow{\underline{\delta(x_1)}} A_1}\\
    ot_3  =& \openrule{
                          \{  \xrightarrow{a_2}_Q \},\\ [\forall y2. a_2\neq acc(y2)] }
    {A_1 \xrightarrow{a_2} A_1}
  \end{array}$
  \end{minipage}
  \hspace{2mm}
  \begin{minipage}{6cm}
$  \begin{array}{@{}r@{}l@{}}
    ot'_1  =& \openrule{
    	\{\xrightarrow{b1}_P\}, ~~ [\forall y1. b1\neq\delta(y1) \!\land\! v=0]}
    {B_0 \xrightarrow{b1} B_0}\\
    ot'_2  =& \openrule{
                            \{\xrightarrow{\delta(y2)}_P,\,
                            \xrightarrow{acc(x2)}_Q \},~
                                        v=0,~
                                        \{v\gets 1\}}
    {B_0 \xrightarrow{\underline{\delta(x2)}} B_1}\\
    ot'_3  =& \openrule{
                            \{\xrightarrow{b2}_Q\}, ~~ [\forall y2. b2\neq\delta(y2) \!\land\! v=1] }
    {B_1 \xrightarrow{b2} B_1}
  \end{array}$
  \end{minipage}
  \caption{The two open automata}
  \label{schema:enable-OAs}
\end{figure}

The open automata generated for the two encodings of Enable in
Figure~\ref{schema:enable-pnets} are displayed in
Figure~\ref{schema:enable-OAs}.
Note that although these automata look very similar to the PLTS
controllers from Figure~\ref{schema:enable-pnets}, their nature is very
different: they represent the behavior of the whole system, including
the holes $P$ and $Q$. Their transitions do not rely on a particular
implementation of the controllers, but only encode the 
relations between the actions of the holes.
We have shown in \cite{henrio:Forte2016} 
\YX{that the two open automata are equivalent with respect to a notion of FH-bisimulation.}
%how to prove that they are bisimilar.

@@

\begin{figure}[h]
  \begin{minipage}{2cm}
  \includegraphics[width=0.9\linewidth]{ActaXFIG/PQR-automaton}
  \end{minipage}
  \hspace{10mm}
  \begin{minipage}{3.5cm}
  \begin{eqnarray*}
    OT_1 & = \openrule{\{\xrightarrow{a_1}_P\}\\ a1\neq\delta(x)}
    {\ostate{00} \xrightarrow{a_1} \ostate{00}}\\
    OT_2 & = \openrule{
                            \{\xrightarrow{\delta(x1)}_P, ~~
                            \xrightarrow{acc(x1)}_Q\}
                      }
                      {\ostate{00} \xrightarrow{\underline{\delta(x1)}} \ostate{10}}    \\
    OT_3 & = \openrule{
                        \{\xrightarrow{a_2}_Q\},\\a2\neq\delta(x)
                        }
    {\ostate{10} \xrightarrow{a_2} \ostate{10}}
    \end{eqnarray*}
  \end{minipage}
  \hspace{10mm}
  \begin{minipage}{4cm}
  \begin{eqnarray*}
    OT_4 & = \openrule{
                      \{\xrightarrow{\delta(x2)}_Q, \qquad 
                       \xrightarrow{acc(x2)}_R\} 
                       }
                      {\ostate{10} \xrightarrow{\underline{\delta(x2)}} \ostate{11}}    \\
    OT_5 & = \openrule{\{\xrightarrow{a_4}_R\}
    	}
    {\ostate{11} \xrightarrow{a_4} \ostate{11}}
    \end{eqnarray*}
  \end{minipage}
  \caption{Open automaton for the term ``$P>>(Q>>R)$''}
  \label{schema:enable3bis}
\end{figure}
